
# Prevalence of visible lesions seperately by SICTT and IFN-γ positive animals at slaughter {#sec-btblesions}

---

$~$ $~$



```{r}
#| label: global-options
#| include: false

knitr::opts_chunk$set(fig.align = "center")
# like this: #| fig-align: center
```

```{r}
#| label: readin-data
#| include: false
#setwd(rprojroot::find_rstudio_root_file())
# read_in_trends_short.R
#source(file.path("R_files", "read_in_quarto_trends.R"))

source("R_files/read_in_quarto_trends.R")
#source("R_files/case_number_quarto_functions.R")

# reactor lesion functions
source("R_files/reactor_lesion_gif_skin_functions.R", echo = T, print.eval = TRUE)


# reading in all_cases_collapsed version from here: gif_lesion.R (called this as the issue was trying to identify which skin cases were gif)
# read in data (from case_lesion_rate / need to rename to reactor_lesion_rate)
# permit <- readRDS("data/edited_output/reactor_lesion_rate.rds")
all_cases_collapsed <- readRDS(paste0("data/edited_output/all_cases_collapsed.rds")) %>% 
    filter(year > 2007) 


# set as TRUE or FALSE
include_co_op_chunks = TRUE

```






\newpage

<!---BLOCK_LANDSCAPE_START--->

## Introduction

This document presents the prevalence of visible lesions at slaughter within bTB positive animals separately by those classified by the SICTT and `r gif_string`. There was no facility within AHCS prior to May 2019 to record `r gif_string` positive animals. Before this, if an animal had a positive `r gif_string` result, their SICTT results were reinterpreted as positive. As a result, we have had to use old raw `r gif_string` data to try and differentiate between which animals were positive on `r gif_string` or SICTT. The old data has some missing tag numbers etc. and so we cannot be certain that we have assigned all cases to either `r gif_string` or SICTT correctly. However, we believe, the vast majority of the data is correct.  **Note, currently, we seem to be missing some lesions for Dec 2023. (awaiting new data dump)** 

This is the difference between overall lesion prevalence document and this one (i.e. we make an attempt to seperate SICTT and `r gif_string` positive trends).

## Definitions

### Proportion of animals with lesions at routine slaughter 

If an animal is classified as bTB positive before entering an abattoir or a meatplant, it must have a permit to enter. Any animal with a permit will get a visual inspection for lesions. In rare instances, a bTB positive animal may be sent to a knackery instead of an abattoir or meatplant (e.g. if the animal dies before being sent to the factory). In these cases, a permit may not be issued and a visual inspection does not routinely take place in a knackery. For these reasons, only animals with a permit (and thus, a visual check completed) are considered for this document. 

Of all SICTT and `r gif_string` positive animals, this is the breakdown of the number of animals with a permit:

```{r}
#| label: fig-permittypes
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 10
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter.")
all_cases_collapsed %>% 
  filter(best_estimate_gif_skin_lab_string != "Lab") %>% 
  tabyl(., per_type)
```


Of importance here is that `ER26GIF` and `ER26SGIF` permits refer to `r gif_string` bTB positive animals. As mentioned above, these did not exist on AHCS prior to May 2019.

$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage


## Overall lesion prevalence (seperately by SICTT or `r gif_string`)

### All animals combined:


```{r}
#| label: fig-overall
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter.")
lesion_rate_by_gif_or_skin_year <- lesion_rate_by_gif_or_skin(all_cases_collapsed, c(year))
reactor_lesion_by_gif_or_skin_plot_fn(lesion_rate_by_gif_or_skin_year, "year", "lesion_rate")
```


$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage


### All animals combined monthly: {.tabset}


$~$ $~$



#### Interactive


```{r}
#| label: fig-overallmonthly
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter.")

lesion_rate_by_gif_or_skin_month <- lesion_rate_by_gif_or_skin(all_cases_collapsed, c(year_month))
min_max_dates_df <- min_max_geom_react(lesion_rate_by_gif_or_skin_month)
reactor_lesion_by_gif_or_skin_plot_fn(lesion_rate_by_gif_or_skin_month, "year_month", "lesion_rate", month_interactive = TRUE)
```







$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage



### AHCS herd-type per year


$~$ $~$


```{r}
#| label: fig-overallahcsyear
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by year.")

lesion_rate_by_gif_or_skin_year_ahcs <- lesion_rate_by_gif_or_skin(all_cases_collapsed, c(year, herd_type))
reactor_lesion_by_gif_or_skin_plot_fn(lesion_rate_by_gif_or_skin_year_ahcs, "year", "lesion_rate", facet_var = herd_type)
```




$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage


### AHCS herd-type per month {.tabset}


$~$ $~$



#### Interactive

$~$


```{r}
#| label: fig-overallahcsmonth
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by month and AHCS herd-type.")


lesion_rate_by_gif_or_skin_month_ahcs <- lesion_rate_by_gif_or_skin(all_cases_collapsed, c(year_month, herd_type))
min_max_dates_df <- min_max_geom_react(lesion_rate_by_gif_or_skin_month_ahcs)
reactor_lesion_by_gif_or_skin_plot_fn(lesion_rate_by_gif_or_skin_month_ahcs, "year_month", "lesion_rate", month_interactive = TRUE, facet_var = herd_type)
```






$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage






### Brock et al. herd-types per year



$~$ $~$


```{r}
#| label: fig-overallbrockyear
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by Brock et al. herd-types per year.")


lesion_rate_by_gif_or_skin_year_brock <- lesion_rate_by_gif_or_skin(all_cases_collapsed, c(year, herd_type_ml_description))
reactor_lesion_by_gif_or_skin_plot_fn(lesion_rate_by_gif_or_skin_year_brock, "year", "lesion_rate", facet_var = herd_type_ml_description)
```

$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage


### Brock et al. herd-types per month {.tabset}


$~$ $~$



#### Interactive

$~$

```{r}
#| label: fig-overallbrockmonth
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by Brock et al. herd-types per month.")


lesion_rate_by_gif_or_skin_month_brock <- lesion_rate_by_gif_or_skin(all_cases_collapsed, c(year_month, herd_type_ml_description))
min_max_dates_df <- min_max_geom_react(lesion_rate_by_gif_or_skin_month_brock)
reactor_lesion_by_gif_or_skin_plot_fn(lesion_rate_by_gif_or_skin_month_brock, "year_month", "lesion_rate", month_interactive = TRUE, facet_var = herd_type_ml_description)
```




$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage

## Lesion prevalence by location

The following section estimates the prevalence of SICTT or `r gif_string` positive animals that have lesions at different locations in the body.


### All animals combined:

Yearly:


```{r}
#| label: fig-lesionlocationyear
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by lesion location per year.")


permit_lesion_head_chest_type_by_gif_skin_year <- permit_lesion_head_chest_type_by_gif_skin(all_cases_collapsed, c(year))
reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_year, "year", "lesion_rate", facet_var = best_estimate_gif_skin_lab_string)
```



$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage

Monthly version:


```{r}
#| label: fig-lesionlocationmonth
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by lesion location per month.")

permit_lesion_head_chest_type_by_gif_skin_month <- permit_lesion_head_chest_type_by_gif_skin(all_cases_collapsed, c(year_month))
min_max_dates_df <- min_max_geom_react(permit_lesion_head_chest_type_by_gif_skin_month)
reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_month, "year_month", "lesion_rate", month_interactive = TRUE, facet_var = best_estimate_gif_skin_lab_string)
```





$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage

### Proportion breakdown by where the lesions are found:

(of animals with lesions, what is the proportion breakdown of where they are located)


SICTT:

```{r}
#| label: fig-proportionskin
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 12 
#| fig.cap: "Proportion of lesions by location for SICTT postive animals by year"

skin_proportion_df <- all_cases_collapsed %>% 
  # drop reactors without permit
  filter(!is.na(per_type)) %>% 
  filter(str_detect(best_estimate_gif_skin_lab_string, "IFN")) %>% 
  filter(!per_type  %in% c(#"ER26B",  (deemed reactors - keeping for now, they were missing test_date but will use per_date instead)
    "ER26D")) %>% 
  filter(!is.na(permit_lesion_type)) %>% 
  filter(permit_lesion_type != "Non-visible lesions") %>% 
  split(.$year) %>%
  map(~tabyl(.x, permit_lesion_type)) %>% 
  map(., ~mutate(.x, percent = round(percent, 2)) %>% select(-n)) %>% 
  imap(~mutate(.x, year = .y)) %>% 
  bind_rows()


#plot barchart (which would be actually values) to compare to ggstream to check if ok?
proportion_skin_lesion_location_barchart <- ggplot(skin_proportion_df, aes(x = year, y = percent, fill = permit_lesion_type)) +
  geom_bar(position = "fill", stat = "identity", color = 'black', width = 0.9) +
  theme_jamie +
  scale_fill_simpsons(name = "Lesion type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  labs(x = "Year", 
       y = "Proportion of lesion type", 
       title = paste0("Proportion of lesions by location for SICTT postive animals"))
       #caption = "we only have data for Brock categorisation available for years 2015-2022")
proportion_skin_lesion_location_barchart
```


$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage

`r gif_string`:

```{r}
#| label: fig-proportiongif
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 12 
#| fig.cap: !expr paste0("Proportion of lesions by location for ", gif_string, " postive animals by year")

gif_proportion_df <- all_cases_collapsed %>% 
  # drop reactors without permit
  filter(!is.na(per_type)) %>% 
  filter(str_detect(best_estimate_gif_skin_lab_string, "IFN")) %>% 
  filter(!per_type  %in% c(#"ER26B",  (deemed reactors - keeping for now, they were missing test_date but will use per_date instead)
    "ER26D")) %>% 
  filter(!is.na(permit_lesion_type)) %>% 
  filter(permit_lesion_type != "Non-visible lesions") %>% 
  split(.$year) %>%
  map(~tabyl(.x, permit_lesion_type)) %>% 
  map(., ~mutate(.x, percent = round(percent, 2)) %>% select(-n)) %>% 
  imap(~mutate(.x, year = .y)) %>% 
  bind_rows()


#plot barchart (which would be actually values) to compare to ggstream to check if ok?
proportion_gif_lesion_location_barchart <- ggplot(gif_proportion_df, aes(x = year, y = percent, fill = permit_lesion_type)) +
  geom_bar(position = "fill", stat = "identity", color = 'black', width = 0.9) +
  theme_jamie +
  scale_fill_simpsons(name = "Lesion type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  labs(x = "Year", 
       y = "Proportion of lesion type", 
       title = paste0("Proportion of lesions by location for ", gif_string, " postive animals"))
       #caption = "we only have data for Brock categorisation available for years 2015-2022")
proportion_gif_lesion_location_barchart
```


$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage


### AHCS herd-types per year


$~$ $~$


$~$

```{r}
#| label: fig-lesionlocationahcsyear
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 12 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by lesion location and AHCS herd-types per year.")

permit_lesion_head_chest_type_by_gif_skin_year_ahcs <- permit_lesion_head_chest_type_by_gif_skin(all_cases_collapsed, c(year, herd_type))

# N.B. rather than doing a 3 way facet - just pull out gif and skin seperate
# reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_year_ahcs %>% 
#                                filter(best_estimate_gif_skin_lab_string == "SICTT"), "year", "lesion_rate", facet_var = herd_type)
# 
# # gif
# reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_year_ahcs %>% 
#                                filter(str_detect(best_estimate_gif_skin_lab_string, 'IFN')), "year", "lesion_rate", facet_var = herd_type)


# it kind of makes more sense to compare herd types by gif and skin
# so by herd type instead of gif/skin above:
# pull out dairy first:
# reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_year_ahcs %>% 
#                                filter(herd_type == "Dairy"), "year", "lesion_rate", facet_var = best_estimate_gif_skin_lab_string)

# ive just edited the facet to fix this
reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_year_ahcs, "year", "lesion_rate", facet_grid_vars = herd_type)
```



$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage




### AHCS herd-types per month {.tabset}


$~$ $~$



#### Interactive

$~$


```{r}
#| label: fig-lesionlocationahcsmonth
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by lesion location and AHCS herd-types per month.")

permit_lesion_head_chest_type_by_gif_skin_month_ahcs <- permit_lesion_head_chest_type_by_gif_skin(all_cases_collapsed, c(year_month, herd_type))
min_max_dates_df <- min_max_geom_react(permit_lesion_head_chest_type_by_gif_skin_month_ahcs)
reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_month_ahcs, "year_month", "lesion_rate", month_interactive = TRUE, facet_grid_vars = herd_type)
```



Same plot but zoomed in for clarity (outliers due to small numbers):


```{r}
#| label: fig-lesionlocationahcsyearzoomed
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by lesion location and AHCS herd-types per month.")

permit_lesion_head_chest_type_by_gif_skin_month_ahcs <- permit_lesion_head_chest_type_by_gif_skin(all_cases_collapsed, c(year_month, herd_type)) 
min_max_dates_df <- min_max_geom_react(permit_lesion_head_chest_type_by_gif_skin_month_ahcs)
reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_month_ahcs, "year_month", "lesion_rate", month_interactive = TRUE, facet_grid_vars = herd_type, 
                             y_zoom = 0.25) 
```





$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage


### Brock et al. herd-types per year


```{r}
#| label: fig-lesionlocationbrockyear
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by lesion location and Brock et al. herd-types per year.")

permit_lesion_head_chest_type_by_gif_skin_year_brock <- permit_lesion_head_chest_type_by_gif_skin(all_cases_collapsed, c(year, herd_type_ml_description)) 
reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_year_brock, "year", "lesion_rate", facet_grid_vars = herd_type_ml_description) 
```


### Brock et al. herd-types per month {.tabset}


$~$ $~$



#### Interactive

$~$



```{r}
#| label: fig-lesionlocationbrockmonth
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by lesion location and Brock et al. herd-types per month.")

permit_lesion_head_chest_type_by_gif_skin_month_brock <- permit_lesion_head_chest_type_by_gif_skin(all_cases_collapsed, c(year_month, herd_type_ml_description)) 
min_max_dates_df <- min_max_geom_react(permit_lesion_head_chest_type_by_gif_skin_month_brock)
reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_month_brock, "year_month", "lesion_rate", month_interactive = TRUE, facet_grid_vars = herd_type_ml_description) 
```


Same plot but zoomed in for clarity (outliers due to small numbers):



```{r}
#| label: fig-lesionlocationbrockmonthzoom
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by lesion location and Brock et al. herd-types per month.")

permit_lesion_head_chest_type_by_gif_skin_month_brock <- permit_lesion_head_chest_type_by_gif_skin(all_cases_collapsed, c(year_month, herd_type_ml_description)) 
min_max_dates_df <- min_max_geom_react(permit_lesion_head_chest_type_by_gif_skin_month_brock)
reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_month_brock, "year_month", "lesion_rate", month_interactive = TRUE, facet_grid_vars = herd_type_ml_description) 
```




## Additional things to add

- Tables
- Smooth trend plots



## Questions

- Why is there a peak in lesions in bTB positive animals at the start of the year (seems more pronounced in SICTT?)? Is this simply an administrative issue whereby there is a delay in sending animals to a factory?

- Would it have anything to do with a slight increase in avian readings around Sept (this is seen in `r gif_string`, we presume the same in SICTT)? The animals have been exposed to the environment for longer at this stage?

- Avian is lower at the start of the year - is this because the majority of animals will be immunosurpressed (pregnant)(`r gif_string` is mainly testing dairy animals)
















