# Prevalence of visible lesions seperately by SICTT and IFN-γ positive animals at slaughter {#sec-btblesions}

------------------------------------------------------------------------

$~$ $~$

```{r}
#| label: global-options
#| include: false

knitr::opts_chunk$set(fig.align = "center")
# like this: #| fig-align: center
```

```{r}
#| label: readin-data
#| include: false
#setwd(rprojroot::find_rstudio_root_file())
# read_in_trends_short.R
#source(file.path("R_files", "read_in_quarto_trends.R"))

source("R_files/read_in_quarto_trends.R")
#source("R_files/case_number_quarto_functions.R")

# reactor lesion functions
source("R_files/reactor_lesion_gif_skin_functions.R", echo = T, print.eval = TRUE)


# reading in all_cases_collapsed version from here: gif_lesion.R (called this as the issue was trying to identify which skin cases were gif)
# read in data (from case_lesion_rate / need to rename to reactor_lesion_rate)
# permit <- readRDS("data/edited_output/reactor_lesion_rate.rds")
# all_cases_collapsed <- readRDS(paste0("data/edited_output/all_cases_collapsed.rds")) %>% 
#     filter(year > 2007) 
all_cases_collapsed_file <- dir_info("data", type = "file", regexp = "(?i)all_cases_collapsed.*\\.rds$") %>% 
  subset(subset = modification_time == max(modification_time), select = path, drop = T)
all_cases_collapsed <- readRDS(all_cases_collapsed_file) %>% 
     filter(year > 2007) %>% 
    filter(year != 2025)

# set as TRUE or FALSE
include_co_op_chunks = TRUE

```

\newpage

<!---BLOCK_LANDSCAPE_START--->

## Introduction

This page presents the prevalence of visible lesions at slaughter within bTB positive animals separately by those classified by the SICTT and `r gif_string`. There was no facility within AHCS prior to May 2019 to record `r gif_string` positive animals. Before this, if an animal had a positive `r gif_string` result, their SICTT results were reinterpreted as positive. As a result, we have had to use old raw `r gif_string` data to try and differentiate between which animals were positive on `r gif_string` or SICTT. The old data has some missing tag numbers etc. and so we cannot be certain that we have assigned all cases to either `r gif_string` or SICTT correctly. However, we believe, the vast majority of the data is correct.

This is the difference between overall lesion prevalence page and this one (i.e. we make an attempt to seperate SICTT and `r gif_string` positive trends).

## Definitions

### Proportion of animals with lesions at routine slaughter

If an animal is classified as bTB positive before entering an abattoir or a meatplant, it must have a permit to enter. Any animal with a permit will get a visual inspection for lesions. In rare instances, a bTB positive animal may be sent to a knackery instead of an abattoir or meatplant (e.g. if the animal dies before being sent to the factory). In these cases, a permit may not be issued and a visual inspection does not routinely take place in a knackery. For these reasons, only animals with a permit (and thus, a visual check completed) are considered for this document.

Of all SICTT and `r gif_string` positive animals, this is the breakdown of the number of animals with a permit:

```{r}
#| label: fig-permittypes
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 10
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter.")
all_cases_collapsed %>% 
  filter(best_estimate_gif_skin_lab_string != "Lab") %>% 
  tabyl(., per_type)
```

Of importance here is that `ER26GIF` and `ER26SGIF` permits refer to `r gif_string` bTB positive animals. As mentioned above, these did not exist on AHCS prior to May 2019.

$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage

## Lesion counts

Before looking at lesion prevalence, the count of lesions by detection method is plotted below as a method to look at absolute numbers which can be telling a different story.


```{r}
#| label: fig-lesioncount
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion counts within SICTT and ", gif_string, " positive animals at slaughter.")
#| 
lesion_count_per_year <- all_cases_collapsed %>% 
  filter(best_estimate_gif_skin_lab  %in% c("skin", "gif")) %>% 
  group_by(year, best_estimate_gif_skin_lab_string) %>% 
  summarise(count_of_lesions = sum(permit_lesions_yes_no, na.rm = TRUE)) %>% 
  ungroup()

#lesion_count_per_year

ggplot(lesion_count_per_year, aes(x = factor(year), y = count_of_lesions, fill = best_estimate_gif_skin_lab_string)) + #color = best_estimate_gif_skin_lab_string)) +
  geom_col() +
  labs(x = "Year",
       y = "Lesion count",
       title = paste0("Lesion count amoung SICTT and ", gif_string, " positive animals"),
       caption = paste0("Definition: ", 
                        "Animals that were classified as bTB positive using ", 
                        "SICTT or IFN-γ and had visible lesions at slaughter")) +
  theme_jamie +
  theme_45_x +
  scale_fill_nejm(name = "Detection type") +
  # scale_y_continuous(labels = scales::label_percent(scale = 100)) +
  # scale_x_type +
  #scale_x_continuous(breaks = nth_element(c(unique(df[[x_var]])), 1, 2)) +
  #scale_color_simpsons(name = "Quartile type") +
  # need this for bold headings
  theme(title = element_text(size = 18, face = "bold"),
        legend.text = element_markdown(size = 16),
        plot.caption.position = "plot",
        legend.title = element_text(size = 16, face = "bold"),
        plot.caption = element_text(size = 9, color = "black", face = "italic"),
        plot.title.position = "plot",
        # remove legend so its zoomed in, then add legend seperately
        legend.position = "top"
        #legend.position = c(0.2, 0.9))  
  ) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 
```






```{r}
#| label: fig-lesioncountherdtype
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion counts within SICTT and ", gif_string, " positive animals at slaughter.")
#| 
brock_cols_named <- c("Dairy" = "#FFB400",
                      "Beef"  = "#3AAFC7",
                      "Store" = "#C20008",
                      "Fattener" = "#009B95",
                      "Mixed"  =  "#25AA82",
                      "Trader" =  "#4BB970",
                      "Unknown" = "#70C85D")
lesion_count_per_year_herd_type <- all_cases_collapsed %>% 
  filter(best_estimate_gif_skin_lab  %in% c("skin", "gif")) %>% 
  group_by(year, herd_type_ml_description) %>% 
  summarise(count_of_lesions = sum(permit_lesions_yes_no, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(!herd_type_ml_description  %in% c("Trader", "Unknown")) %>% 
  droplevels() %>% 
  filter(!is.na(herd_type_ml_description))


ggplot(lesion_count_per_year_herd_type, aes(x = factor(year), y = count_of_lesions, fill = herd_type_ml_description)) + #color = best_estimate_gif_skin_lab_string)) +
  geom_col(position = position_dodge2(preserve = "single", padding = 0.1)) +
  labs(x = "Year",
       y = "Lesion count",
       title = paste0("Lesion count amoung SICTT and ", gif_string, " positive animals"),
       caption = paste0("Definition: ", 
                        "Animals that were classified as bTB positive using ", 
                        "SICTT or IFN-γ and had visible lesions at slaughter")) +
  theme_jamie +
  theme_45_x +
  scale_fill_manual(name = "Herd-type", values = brock_cols_named) +
  # scale_y_continuous(labels = scales::label_percent(scale = 100)) +
  # scale_x_type +
  #scale_x_continuous(breaks = nth_element(c(unique(df[[x_var]])), 1, 2)) +
  #scale_color_simpsons(name = "Quartile type") +
  # need this for bold headings
  theme(title = element_text(size = 18, face = "bold"),
        legend.text = element_markdown(size = 16),
        plot.caption.position = "plot",
        legend.title = element_text(size = 16, face = "bold"),
        plot.caption = element_text(size = 9, color = "black", face = "italic"),
        plot.title.position = "plot",
        # remove legend so its zoomed in, then add legend seperately
        legend.position = "top"
        #legend.position = c(0.2, 0.9))  
  ) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 
```


## Overall lesion prevalence (seperately by SICTT or `r gif_string`)

### All animals combined:

```{r}
#| label: fig-overall
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter.")
lesion_rate_by_gif_or_skin_year <- lesion_rate_by_gif_or_skin(all_cases_collapsed, c(year))
reactor_lesion_by_gif_or_skin_plot_fn(lesion_rate_by_gif_or_skin_year, "year", "lesion_rate")
```

$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage

### All animals combined monthly: {.tabset}

$~$ $~$

#### Interactive

```{r}
#| label: fig-overallmonthly
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter.")

lesion_rate_by_gif_or_skin_month <- lesion_rate_by_gif_or_skin(all_cases_collapsed, c(year_month))
min_max_dates_df <- min_max_geom_react(lesion_rate_by_gif_or_skin_month)
reactor_lesion_by_gif_or_skin_plot_fn(lesion_rate_by_gif_or_skin_month, "year_month", "lesion_rate", month_interactive = TRUE)
```

$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage

### AHCS herd-type per year

$~$ $~$

```{r}
#| label: fig-overallahcsyear
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by year.")

lesion_rate_by_gif_or_skin_year_ahcs <- lesion_rate_by_gif_or_skin(all_cases_collapsed, c(year, herd_type))
reactor_lesion_by_gif_or_skin_plot_fn(lesion_rate_by_gif_or_skin_year_ahcs, "year", "lesion_rate", facet_var = herd_type)
```

$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage

### AHCS herd-type per month {.tabset}

$~$ $~$

#### Interactive

$~$

```{r}
#| label: fig-overallahcsmonth
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by month and AHCS herd-type.")


lesion_rate_by_gif_or_skin_month_ahcs <- lesion_rate_by_gif_or_skin(all_cases_collapsed, c(year_month, herd_type))
min_max_dates_df <- min_max_geom_react(lesion_rate_by_gif_or_skin_month_ahcs)
reactor_lesion_by_gif_or_skin_plot_fn(lesion_rate_by_gif_or_skin_month_ahcs, "year_month", "lesion_rate", month_interactive = TRUE, facet_var = herd_type)
```

$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage

### Brock et al. herd-types per year

$~$ $~$

```{r}
#| label: fig-overallbrockyear
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by Brock et al. herd-types per year.")


lesion_rate_by_gif_or_skin_year_brock <- lesion_rate_by_gif_or_skin(all_cases_collapsed, c(year, herd_type_ml_description))
reactor_lesion_by_gif_or_skin_plot_fn(lesion_rate_by_gif_or_skin_year_brock %>% 
                                        filter(herd_type_ml_description != "Store"), "year", "lesion_rate", facet_var = herd_type_ml_description)
```

$~$ $~$

------------------------------------------------------------------------

$~$ $~$


#### Overall lesion rate in both combined

$~$ $~$

```{r}
#| label: fig-overallbrockyearcombined
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by Brock et al. herd-types per year.")


lesion_rate_overall_cases_year = lesion_rate_by_gif_or_skin_year_brock %>% 
  group_by(year, herd_type_ml_description) %>% 
  summarise(count_of_lesions = sum(count_of_lesions),
            total_gif_or_skin_cases = sum(total_gif_or_skin_cases),
            year_f = first(year_f)) %>% 
  ungroup() %>% 
  mutate(lesion_rate = count_of_lesions/total_gif_or_skin_cases)

lesion_rate_overall_cases_year <- lesion_rate_overall_cases_year %>% 
  mutate(herd_type_ml_description = fct_relevel(herd_type_ml_description, c("Dairy", "Beef", "Store", "Fattener",
                                                                            "Mixed", "Trader",
                                                                            #"Seasonal", now removed
                                                                            "Unknown")))
#levels(lesion_rate_overall_cases_year$herd_type_ml_description)


# remove unknown herd type (although these are interesting in themselves)
lesion_rate_overall_cases_year_plot <- ggplot(lesion_rate_overall_cases_year %>% 
                                                filter(herd_type_ml_description != "Unknown") %>% 
                                                filter(herd_type_ml_description != "Trader") %>% 
                                                droplevels(), aes(x = year, y = lesion_rate, color = herd_type_ml_description)) +
  geom_line(linewidth = 2) +
  labs(x = "Year", 
       y = "Lesion prevalence", 
       title = "Lesion prevalence within bTB positive animals classified by SICTT or IFN-γ",
       caption = paste0(all_lesions_vector,
                        "\n",
                        "Unknown and Trader herds excluded due to low numbers")) +
  theme_jamie +
  scale_color_manual(name = "Brock et al. herd type", values = brock_cols) +
  scale_y_continuous(labels = scales::label_percent(scale = 100)) +
  scale_x_continuous(breaks = nth_element(c(unique(lesion_rate_overall_cases_year$year)), 1, 2)) +
  theme(title = element_text(size = 18, face = "bold"),
        legend.text = element_markdown(size = 16),
        plot.caption.position = "plot",
        plot.title.position = "plot",
        
        legend.title = element_text(size = 16, face = "bold"),
        plot.caption = element_text(size = 9, color = "black", face = "italic"),
  )
lesion_rate_overall_cases_year_plot
```



\newpage

### Brock et al. herd-types per month {.tabset}

$~$ $~$

#### Interactive

$~$

```{r}
#| label: fig-overallbrockmonth
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by Brock et al. herd-types per month.")


lesion_rate_by_gif_or_skin_month_brock <- lesion_rate_by_gif_or_skin(all_cases_collapsed, c(year_month, herd_type_ml_description))
min_max_dates_df <- min_max_geom_react(lesion_rate_by_gif_or_skin_month_brock)
reactor_lesion_by_gif_or_skin_plot_fn(lesion_rate_by_gif_or_skin_month_brock, "year_month", "lesion_rate", month_interactive = TRUE, facet_var = herd_type_ml_description)
```

$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage

## Lesion prevalence by location

The following section estimates the prevalence of SICTT or `r gif_string` positive animals that have lesions at different locations in the body.

### All animals combined:

Yearly:

```{r}
#| label: fig-lesionlocationyear
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by lesion location per year.")


permit_lesion_head_chest_type_by_gif_skin_year <- permit_lesion_head_chest_type_by_gif_skin(all_cases_collapsed, c(year))
reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_year, "year", "lesion_rate", facet_var = best_estimate_gif_skin_lab_string)
```

$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage

Monthly version:

```{r}
#| label: fig-lesionlocationmonth
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by lesion location per month.")

permit_lesion_head_chest_type_by_gif_skin_month <- permit_lesion_head_chest_type_by_gif_skin(all_cases_collapsed, c(year_month))
min_max_dates_df <- min_max_geom_react(permit_lesion_head_chest_type_by_gif_skin_month)
reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_month, "year_month", "lesion_rate", month_interactive = TRUE, facet_var = best_estimate_gif_skin_lab_string)
```

$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage

### Proportion breakdown by where the lesions are found:

(of animals with lesions, what is the proportion breakdown of where they are located)

SICTT:

```{r}
#| label: fig-proportionskin
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 12 
#| fig.cap: "Proportion of lesions by location for SICTT postive animals by year"

skin_proportion_df <- all_cases_collapsed %>% 
  # drop reactors without permit
  filter(!is.na(per_type)) %>% 
  filter(str_detect(best_estimate_gif_skin_lab_string, "IFN")) %>% 
  filter(!per_type  %in% c(#"ER26B",  (deemed reactors - keeping for now, they were missing test_date but will use per_date instead)
    "ER26D")) %>% 
  filter(!is.na(permit_lesion_type)) %>% 
  filter(permit_lesion_type != "Non-visible lesions") %>% 
  split(.$year) %>%
  map(~tabyl(.x, permit_lesion_type)) %>% 
  map(., ~mutate(.x, percent = round(percent, 2)) %>% select(-n)) %>% 
  imap(~mutate(.x, year = .y)) %>% 
  bind_rows()


#plot barchart (which would be actually values) to compare to ggstream to check if ok?
proportion_skin_lesion_location_barchart <- ggplot(skin_proportion_df, aes(x = year, y = percent, fill = permit_lesion_type)) +
  geom_bar(position = "fill", stat = "identity", color = 'black', width = 0.9) +
  theme_jamie +
  scale_fill_simpsons(name = "Lesion type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  labs(x = "Year", 
       y = "Proportion of lesion type", 
       title = paste0("Proportion of lesions by location for SICTT postive animals"))
       #caption = "we only have data for Brock categorisation available for years 2015-2022")
proportion_skin_lesion_location_barchart
```

$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage

`r gif_string`:

```{r}
#| label: fig-proportiongif
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 12 
#| fig.cap: !expr paste0("Proportion of lesions by location for ", gif_string, " postive animals by year")

gif_proportion_df <- all_cases_collapsed %>% 
  # drop reactors without permit
  filter(!is.na(per_type)) %>% 
  filter(str_detect(best_estimate_gif_skin_lab_string, "IFN")) %>% 
  filter(!per_type  %in% c(#"ER26B",  (deemed reactors - keeping for now, they were missing test_date but will use per_date instead)
    "ER26D")) %>% 
  filter(!is.na(permit_lesion_type)) %>% 
  filter(permit_lesion_type != "Non-visible lesions") %>% 
  split(.$year) %>%
  map(~tabyl(.x, permit_lesion_type)) %>% 
  map(., ~mutate(.x, percent = round(percent, 2)) %>% select(-n)) %>% 
  imap(~mutate(.x, year = .y)) %>% 
  bind_rows()


#plot barchart (which would be actually values) to compare to ggstream to check if ok?
proportion_gif_lesion_location_barchart <- ggplot(gif_proportion_df, aes(x = year, y = percent, fill = permit_lesion_type)) +
  geom_bar(position = "fill", stat = "identity", color = 'black', width = 0.9) +
  theme_jamie +
  scale_fill_simpsons(name = "Lesion type") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  labs(x = "Year", 
       y = "Proportion of lesion type", 
       title = paste0("Proportion of lesions by location for ", gif_string, " postive animals"))
       #caption = "we only have data for Brock categorisation available for years 2015-2022")
proportion_gif_lesion_location_barchart
```

$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage

### AHCS herd-types per year

$~$ $~$

$~$

```{r}
#| label: fig-lesionlocationahcsyear
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 12 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by lesion location and AHCS herd-types per year.")

permit_lesion_head_chest_type_by_gif_skin_year_ahcs <- permit_lesion_head_chest_type_by_gif_skin(all_cases_collapsed, c(year, herd_type))

# N.B. rather than doing a 3 way facet - just pull out gif and skin seperate
# reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_year_ahcs %>% 
#                                filter(best_estimate_gif_skin_lab_string == "SICTT"), "year", "lesion_rate", facet_var = herd_type)
# 
# # gif
# reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_year_ahcs %>% 
#                                filter(str_detect(best_estimate_gif_skin_lab_string, 'IFN')), "year", "lesion_rate", facet_var = herd_type)


# it kind of makes more sense to compare herd types by gif and skin
# so by herd type instead of gif/skin above:
# pull out dairy first:
# reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_year_ahcs %>% 
#                                filter(herd_type == "Dairy"), "year", "lesion_rate", facet_var = best_estimate_gif_skin_lab_string)

# ive just edited the facet to fix this
reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_year_ahcs, "year", "lesion_rate", facet_grid_vars = herd_type)
```

$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage

### AHCS herd-types per month {.tabset}

$~$ $~$

#### Interactive

$~$

```{r}
#| label: fig-lesionlocationahcsmonth
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by lesion location and AHCS herd-types per month.")

permit_lesion_head_chest_type_by_gif_skin_month_ahcs <- permit_lesion_head_chest_type_by_gif_skin(all_cases_collapsed, c(year_month, herd_type))
min_max_dates_df <- min_max_geom_react(permit_lesion_head_chest_type_by_gif_skin_month_ahcs)
reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_month_ahcs, "year_month", "lesion_rate", month_interactive = TRUE, facet_grid_vars = herd_type)
```

Same plot but zoomed in for clarity (outliers due to small numbers):

```{r}
#| label: fig-lesionlocationahcsyearzoomed
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by lesion location and AHCS herd-types per month.")

permit_lesion_head_chest_type_by_gif_skin_month_ahcs <- permit_lesion_head_chest_type_by_gif_skin(all_cases_collapsed, c(year_month, herd_type)) 
min_max_dates_df <- min_max_geom_react(permit_lesion_head_chest_type_by_gif_skin_month_ahcs)
reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_month_ahcs, "year_month", "lesion_rate", month_interactive = TRUE, facet_grid_vars = herd_type, 
                             y_zoom = 0.25) 
```

$~$ $~$

------------------------------------------------------------------------

$~$ $~$

\newpage

### Brock et al. herd-types per year

```{r}
#| label: fig-lesionlocationbrockyear
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by lesion location and Brock et al. herd-types per year.")

permit_lesion_head_chest_type_by_gif_skin_year_brock <- permit_lesion_head_chest_type_by_gif_skin(all_cases_collapsed, c(year, herd_type_ml_description)) 
reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_year_brock, "year", "lesion_rate", facet_grid_vars = herd_type_ml_description) 
```

### Brock et al. herd-types per month {.tabset}

$~$ $~$

#### Interactive

$~$

```{r}
#| label: fig-lesionlocationbrockmonth
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by lesion location and Brock et al. herd-types per month.")

permit_lesion_head_chest_type_by_gif_skin_month_brock <- permit_lesion_head_chest_type_by_gif_skin(all_cases_collapsed, c(year_month, herd_type_ml_description)) 
min_max_dates_df <- min_max_geom_react(permit_lesion_head_chest_type_by_gif_skin_month_brock)
reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_month_brock, "year_month", "lesion_rate", month_interactive = TRUE, facet_grid_vars = herd_type_ml_description) 
```

Same plot but zoomed in for clarity (outliers due to small numbers):

```{r}
#| label: fig-lesionlocationbrockmonthzoom
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: !expr paste0("Overall lesion prevalence within SICTT and ", gif_string, " positive animals at slaughter by lesion location and Brock et al. herd-types per month.")

permit_lesion_head_chest_type_by_gif_skin_month_brock <- permit_lesion_head_chest_type_by_gif_skin(all_cases_collapsed, c(year_month, herd_type_ml_description)) 
min_max_dates_df <- min_max_geom_react(permit_lesion_head_chest_type_by_gif_skin_month_brock)
reactor_gif_or_skin_chest_fn(permit_lesion_head_chest_type_by_gif_skin_month_brock, "year_month", "lesion_rate", month_interactive = TRUE, facet_grid_vars = herd_type_ml_description) 
```

## Additional things to add

-   Tables
-   Smooth trend plots

## Questions

-   Why is there a peak in lesions in bTB positive animals at the start of the year (seems more pronounced in SICTT?)? Is this simply an administrative issue whereby there is a delay in sending animals to a factory?

-   Would it have anything to do with a slight increase in avian readings around Sept (this is seen in `r gif_string`, we presume the same in SICTT)? The animals have been exposed to the environment for longer at this stage?

-   Avian is lower at the start of the year - is this because the majority of animals will be immunosurpressed (pregnant)(`r gif_string` is mainly testing dairy animals)
