# SICTT test-types {#sec-testtypes}

------------------------------------------------------------------------

```{r}
#| label: global-options
#| include: false

# Center figures (alternative approach in Quarto)
knitr::opts_chunk$set(fig.align = "center")
# like this: #| fig-align: center
```

```{r}
#| label: readin-data
#| include: false
#setwd(rprojroot::find_rstudio_root_file())
# read_in_trends_short.R
#source(file.path("R_files", "read_in_quarto_trends.R"))

source("R_files/read_in_quarto_trends.R")
source("R_files/apt_functions.R")
bd_df_file <- dir_info("data", type = "file", regexp = "(?i)bd_df.*\\.rds$") %>% 
  subset(subset = modification_time == max(modification_time), select = path, drop = T)
#bd_df <- readRDS("data/bd_df_14_Nov_2024.rds")
bd_df <- readRDS(bd_df_file) %>% 
  filter(bd_start_yr > 2007 & bd_start_yr < 2025)
bd_df_labels <- as.list(sjlabelled::get_label(bd_df))
bd_df <- sjlabelled::remove_all_labels(bd_df)

# master_tb
#master_tb <- readRDS("data/master_tb_14_Nov_2024.rds")
master_file <- dir_info("data", type = "file", regexp = "(?i)master.*\\.rds$") %>% 
  subset(subset = modification_time == max(modification_time), select = path, drop = T)
master_tb <- readRDS(master_file) %>% 
  filter(skin_fixed_test_date_year > 2007) %>% 
  filter(skin_fixed_test_date_year != 2025) %>% 
    mutate(test_type_description_number = paste0(test_type, "; ", test_type_description),
         #contig_test_type_binary = ifelse(test_type == "8", 1, 0),
         #contig_test_type_character = ifelse(contig_test_type_binary == 0, "All other test types", "Contig test type")
         ) %>% 
  mutate(test_type_animal_or_herd = case_when(test_type  %in% c("1", "2", "9A", "10A", "8", "7B",
                                                                "5A", "5F", "5C", "5E", "5B", 
                                                                "4", "10B", "10C", "10D",
                                                                "7A") ~ "Herd",
                                              test_type  %in% c("3", "6", "5G", "5H", "5D") ~ "Animal"))



years_for_study <- unique(master_tb$skin_fixed_test_date_year)
# we want dairy at the bottom so we can see how much its going
# order_herd_type <- levels(factor(master_tb$herd_type_ml_sub_class_description))
# order_herd_type
# better_order_herd_type <- c(str_subset(order_herd_type, "^Dairy"), 
#                             str_subset(order_herd_type, "^Beef"),
#                             str_subset(order_herd_type, "^Store"),
#                             "Fattener", "Mixed", "Trader", 
#                             #"Seasonal", #"Rearing dairy females", 
#                             "Unknown")
# better_order_herd_type
# 
# master_tb <- master_tb %>% 
#   mutate(herd_type_ml_description = fct_relevel(herd_type_ml_description, c("Dairy", "Beef", "Store", "Fattener",
#                                                                             "Mixed", "Trader", 
#                                                                             #"Seasonal", now removed
#                                                                             "Unknown")),
#          herd_type_ml_sub_class_description = fct_relevel(herd_type_ml_sub_class_description, better_order_herd_type))
#levels(master_tb$herd_type_ml_description)
#levels(master_tb$herd_type_ml_sub_class_description)
#tabyl(master_tb, herd_type_ml_description)
#tabyl(master_tb, herd_type_ml_sub_class_description)


```

$~$ $~$

## All test-types (herd and animal-level SICTTs)

```{r}
#| label: fig-propsicttalltests
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 9
#| fig.cap: "The proportion of all SICTTs by all test-types."



## all test types ----
all_test_types_df_year <- master_tb %>% 
  # 5D; Factory return (Animal/herd) obsolete
  filter(test_type != "5D") %>% 
  split(list(.$skin_fixed_test_date_year)) %>%
  map(~tabyl(.x, test_type_description_number)) %>% 
  map(., ~mutate(.x, percent = round(percent, 2)) %>% select(-n)) %>% 
  imap(~mutate(.x, year = .y)) %>% 
  bind_rows() %>%
  mutate(year = as.numeric(year))
#all_test_types_df_year


all_test_types_year_bar_plot <- ggplot(all_test_types_df_year, aes(x = year, y = percent, fill = test_type_description_number)) +
  geom_bar(position = position_fill(reverse = TRUE), stat = "identity", #color = 'black', 
           width = 0.9) +
  scale_fill_tableau20(name = "SICTT test-type", direction = 1) +
  #scale_fill_viridis_d(name = "SICTT test-type", direction = 1) +
  theme_cowplot() +
  theme(axis.text = element_text(size = 12, face = "bold", colour = "black"),
        axis.title = element_text(size = 16, face = "bold", colour = "black"),
        #axis.text.x = element_blank(), #remove x axis title
        #axis.text.x = element_text(size = 12)
        #legend title only needed for first one as only extracting first
        legend.title = element_text(size = 10, face = "bold"), 
        legend.text = element_text(size = 10),
        #legend.position = "top",
        #grey background like here:
        #https://github.com/davidsjoberg/ggstream
        #plot.background = element_rect(fill = "grey88", color = NA),
        #make facet white rather than default grey
        #strip.background = element_rect(colour = "white", fill = "white")
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  ) + 
  background_grid() +
  labs(x = "Year", 
       y = "Proportion of all SICTT ", 
       #fill = "AHCS herd type",
       title =
         paste0("Proportion of all SICTT test-types by year"),
       caption = "Test-type 5D; Factory return (Animal/herd) obsolete, removed") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  scale_x_continuous(breaks = years_for_study) +
  scale_y_continuous(labels = scales::label_percent(scale = 100)) 

all_test_types_year_bar_plot
```

------------------------------------------------------------------------

$~$ $~$

### Herd vs animal level test-types

Many SICTT are just animal-level tests and are not associated with a full herd test (e.g., a pre-movement test). We assigned each test-type to a herd or animal level test:

```{r}
#| label: tbl-tableherdanimaltesttypes
#| tbl-cap: "SICTT by herd or animal-level test."
#| include: TRUE
#| echo: FALSE
#| eval: true
#| warning: FALSE
#| message: FALSE
test_type_by_animal_herd <- master_tb %>% 
  select(test_type_description_number, test_type_animal_or_herd) %>% 
  group_by(test_type_description_number) %>% 
  summarise(test_type_animal_or_herd = first(test_type_animal_or_herd)) %>% 
  ungroup() %>% 
  arrange(str_rank(test_type_description_number, numeric = TRUE)) %>% 
  group_by(test_type_animal_or_herd) %>%
  mutate(row = row_number()) %>%
  pivot_wider(
    names_from = test_type_animal_or_herd,
    values_from = test_type_description_number
  ) %>%
  select(row, Herd, Animal) %>%
  arrange(row) %>% 
  select(-row) 

ft <- flextable(test_type_by_animal_herd) %>% 
  align_text_col(align = "center") %>% 
  align(i = NULL, j = 1, align = "left", part = "body") %>% 
  bold(part = "header") %>% 
  #flextable::set_header_labels(values = header_names) %>% # bold header (split up due to weird error)
  #bold(j = c(1, 2)) %>% 
  set_caption(paste0("SICTT by herd or animal-level test.")) %>% 
  #bold(i = 1) %>%  # bold header
  #add_footer_row(values = glue::glue("AHCS: Animal Health Computer System;\nNo. of herds: count of unique herd numbers that had at least one SICTT or IFN-γ \nControlled finishing units (CFUs): only have data up to 2023 \nAHCS herd type: herd type that is assigned to the herd administratively on AHCS \nPopulation figures are based on the number of animals which were in each herd type (Brock et al.) on 1st of January, May, September of each year using the Animal Identification and Movement System (AIM) \nProportion values represent the proportion of herds within each herd classifications within a given year (i.e., vertical) whereas in the case of CFUs, they represent the proportion in each year (i.e., horizontal)"),
  #               colwidths = length(test_type_by_animal_herd)) %>% 
  #width of column if too narrow (only visable in word, not )
  width(width = 2) %>% 
  fontsize(part = "footer", size = 10) %>% 
  #title autonum = autonum numbering in word if needed
  autofit() %>% 
  set_table_properties(layout = "autofit")

ft
```

$~$ $~$

```{r}
#| label: fig-propherdanimaltest
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 9
#| fig.cap: "The proportion of all SICTTs by whether it was an animal or herd-level SICTT."


test_type_animal_or_herd_df_year <- master_tb %>% 
  # 5D; Factory return (Animal/herd) obsolete
  #filter(test_type != "5D") %>% 
  split(list(.$skin_fixed_test_date_year)) %>%
  map(~tabyl(.x, test_type_animal_or_herd)) %>% 
  map(., ~mutate(.x, percent = round(percent, 2)) %>% select(-n)) %>% 
  imap(~mutate(.x, year = .y)) %>% 
  bind_rows() %>%
  mutate(year = as.numeric(year))
#test_type_animal_or_herd_df_year


test_type_animal_or_herd_year_bar_plot <- ggplot(test_type_animal_or_herd_df_year, aes(x = year, y = percent, fill = test_type_animal_or_herd)) +
  geom_bar(position = position_fill(reverse = TRUE), stat = "identity", #color = 'black', 
           width = 0.9) +
  scale_fill_tableau20(name = "SICTT test-type", direction = 1) +
  #scale_fill_viridis_d(name = "SICTT test-type", direction = 1) +
  theme_cowplot() +
  theme(axis.text = element_text(size = 12, face = "bold", colour = "black"),
        axis.title = element_text(size = 16, face = "bold", colour = "black"),
        #axis.text.x = element_blank(), #remove x axis title
        #axis.text.x = element_text(size = 12)
        #legend title only needed for first one as only extracting first
        legend.title = element_text(size = 10, face = "bold"), 
        legend.text = element_text(size = 10),
        legend.position = "top",
        #grey background like here:
        #https://github.com/davidsjoberg/ggstream
        #plot.background = element_rect(fill = "grey88", color = NA),
        #make facet white rather than default grey
        #strip.background = element_rect(colour = "white", fill = "white")
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  ) + 
  background_grid() +
  labs(x = "Year", 
       y = "Proportion of all SICTT ", 
       #fill = "AHCS herd type",
       title =
         paste0("Proportion of all SICTT by whether herd or animal level test")) +
       #caption = "Test-type 5D; Factory return (Animal/herd) obsolete, removed") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  scale_x_continuous(breaks = years_for_study) +
  scale_y_continuous(labels = scales::label_percent(scale = 100)) 

test_type_animal_or_herd_year_bar_plot
```

------------------------------------------------------------------------

$~$ $~$

repeat above plots by herd-level only SICTTs (increase in animal level tests skewing true )

### All test-types (herd-level SICTT only)

```{r}
#| label: fig-propsicttalltestsherdonly
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 11
#| fig.cap: "The proportion of all SICTTs by all test-types."



all_test_types_df_year <- master_tb %>% 
  filter(test_type_animal_or_herd == "Herd") %>% 
  # 5D; Factory return (Animal/herd) obsolete
  filter(test_type != "5D") %>% 
  split(list(.$skin_fixed_test_date_year)) %>%
  map(~tabyl(.x, test_type_description_number)) %>% 
  map(., ~mutate(.x, percent = round(percent, 2)) %>% select(-n)) %>% 
  imap(~mutate(.x, year = .y)) %>% 
  bind_rows() %>%
  mutate(year = as.numeric(year))
#all_test_types_df_year


all_test_types_year_bar_plot <- ggplot(all_test_types_df_year, aes(x = year, y = percent, fill = test_type_description_number)) +
  geom_bar(position = position_fill(reverse = TRUE), stat = "identity", #color = 'black', 
           width = 0.9) +
  scale_fill_tableau20(name = "SICTT test-type", direction = 1) +
  #scale_fill_viridis_d(name = "SICTT test-type", direction = 1) +
  theme_cowplot() +
  theme(axis.text = element_text(size = 12, face = "bold", colour = "black"),
        axis.title = element_text(size = 16, face = "bold", colour = "black"),
        #axis.text.x = element_blank(), #remove x axis title
        #axis.text.x = element_text(size = 12)
        #legend title only needed for first one as only extracting first
        legend.title = element_text(size = 10, face = "bold"), 
        legend.text = element_text(size = 10),
        #legend.position = "top",
        #grey background like here:
        #https://github.com/davidsjoberg/ggstream
        #plot.background = element_rect(fill = "grey88", color = NA),
        #make facet white rather than default grey
        #strip.background = element_rect(colour = "white", fill = "white")
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  ) + 
  background_grid() +
  labs(x = "Year", 
       y = "Proportion of all SICTT ", 
       #fill = "AHCS herd type",
       title =
         paste0("Proportion of all SICTT test-types by year"),
       caption = "Test-type 5D; Factory return (Animal/herd) obsolete, removed") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  scale_x_continuous(breaks = years_for_study) +
  scale_y_continuous(labels = scales::label_percent(scale = 100)) 

all_test_types_year_bar_plot
```

And table format:

```{r}
#| label: tbl-tableherdlevelsictt
#| tbl-cap: "Proportion of herd-level SICTTs by all herd-level test-types (presented as %)"
#| include: TRUE
#| echo: FALSE
#| eval: true
#| warning: FALSE
#| message: FALSE
all_test_types_df_year_tab <- master_tb %>% 
  filter(test_type_animal_or_herd == "Herd") %>% 
  # 5D; Factory return (Animal/herd) obsolete
  filter(test_type != "5D") %>% 
  split(list(.$skin_fixed_test_date_year)) %>%
  map(~tabyl(.x, test_type_description_number)) %>% 
  map(., ~mutate(.x, percent = round(percent, 3)*100) %>% select(-n)) %>% 
  imap(~set_names(.x, ifelse(names(.x) != "test_type_description_number", paste(.y, names(.x), sep = "_"), names(.x)))) %>%
  reduce(full_join, by = "test_type_description_number")

header_names <- setNames(c("Test type", min(master_tb$skin_fixed_test_date_year):max(master_tb$skin_fixed_test_date_year)),
                         names(all_test_types_df_year_tab))
ft <- flextable(all_test_types_df_year_tab) %>% 
  align_text_col(align = "center") %>% 
  align(i = NULL, j = 1, align = "left", part = "body") %>% 
  bold(part = "header") %>% 
  flextable::set_header_labels(values = header_names) %>% # bold header (split up due to weird error)
  bold(j = c(1)) %>% 
  set_caption(paste0("Proportion of herd-level SICTTs by all herd-level test-types (presented as %)")) %>% 
  #bold(i = 1) %>%  # bold header
  # add_footer_row(values = glue::glue("AHCS: Animal Health Computer System;\nNo. of herds: count of unique herd numbers that had at least one SICTT or IFN-γ \nControlled finishing units (CFUs): only have data up to 2023 \nAHCS herd type: herd type that is assigned to the herd administratively on AHCS \nPopulation figures are based on the number of animals which were in each herd type (Brock et al.) on 1st of January, May, September of each year using the Animal Identification and Movement System (AIM) \nProportion values represent the proportion of herds within each herd classifications within a given year (i.e., vertical) whereas in the case of CFUs, they represent the proportion in each year (i.e., horizontal)"),
  #                colwidths = length(final_table_herd_demographics)) %>% 
  #width of column if too narrow (only visable in word, not )
  width(width = 2) %>% 
  fontsize(part = "footer", size = 10) %>% 
  #title autonum = autonum numbering in word if needed
  autofit() %>% 
  set_table_properties(layout = "autofit")

ft
```

## Timing of round test-type

**Seasonality of annual round test (test_type = 1)**

Is there more annual round test during certain months of the year (e.g., summer/autumn) which might reflect more cases being detected?

Simple count and proportion is plotted below.

### All herds count

```{r}
#| label: fig-countsicttroundtest
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 11
#| fig.cap: "The count of all round tests by month"
library(viridis)
round_tests_count <- master_tb %>% 
  filter(test_type == "1") %>% 
  mutate(skin_month = lubridate::month(skin_fixed_test_date, label = TRUE)) %>% 
  select(skin_fixed_test_date_year, skin_month) %>% 
  group_by(skin_fixed_test_date_year, skin_month) %>% 
  summarise(count_of_round_tests_per_month = n()) %>% 
  ungroup() 
# round_tests_count$skin_fixed_test_date_year <- as.numeric(as.character(round_tests$skin_fixed_test_date_year))
# round_tests$skin_fixed_test_date_year <- as.integer(as.character(round_tests$skin_fixed_test_date_year))


ggplot(round_tests_count, aes(x = skin_month, y = count_of_round_tests_per_month)) +
  geom_line(aes(color = skin_fixed_test_date_year, group = skin_fixed_test_date_year), size = 1) +
  scale_color_viridis_c(
    option = "plasma",  # You can try "viridis", "inferno", "magma", etc.
    name = "Year"
  ) +
  labs(
    x = "Month",
    y = "Number of annual round tests (test type = 1)",
    color = "Year",
    title = "Monthly count of annual round tests (test type = 1) by year"
  ) +
  theme_jamie +
  theme_45_x +
  theme(plot.caption.position = "plot",
        #legend.title = element_text(size = 16, face = "bold"),
        plot.caption = element_text(size = 9, color = "black", face = "italic"),
        plot.title.position = "plot")
```

### Count by herd-type

::: panel-tabset
## Fixed axis (for comparison)

```{r}
#| label: fig-countsicttroundtestherdtype
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 11
#| fig.cap: "The count of all round tests by month"
round_tests_count_herd_type <- master_tb %>% 
  filter(!is.na(herd_type_ml_description)) %>% 
  filter(test_type == "1") %>% 
  mutate(skin_month = lubridate::month(skin_fixed_test_date, label = TRUE)) %>% 
  select(skin_fixed_test_date_year, skin_month, herd_type_ml_description) %>% 
  group_by(skin_fixed_test_date_year, skin_month, herd_type_ml_description) %>% 
  summarise(count_of_round_tests_per_month = n()) %>% 
  ungroup() 
#round_tests_count_herd_type


ggplot(round_tests_count_herd_type, aes(x = skin_month, y = count_of_round_tests_per_month)) +
  geom_line(aes(color = skin_fixed_test_date_year, group = skin_fixed_test_date_year), size = 1) +
  scale_color_viridis_c(
    option = "plasma",  # You can try "viridis", "inferno", "magma", etc.
    name = "Year"
  ) +
  labs(
    x = "Month",
    y = "Number of annual round tests (test type = 1)",
    color = "Year",
    title = "Monthly count of annual round tests (test type = 1) by year"
  ) +
  theme_jamie +
  theme_45_x +
  theme(plot.caption.position = "plot",
        #legend.title = element_text(size = 16, face = "bold"),
        plot.caption = element_text(size = 9, color = "black", face = "italic"),
        plot.title.position = "plot") +
  facet_wrap(~herd_type_ml_description, ncol = 2)
```

## Free axis

```{r}
#| label: fig-countsicttroundtestherdtypefreeaxis
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 11
#| fig.cap: "The count of all round tests by month"
ggplot(round_tests_count_herd_type, aes(x = skin_month, y = count_of_round_tests_per_month)) +
  geom_line(aes(color = skin_fixed_test_date_year, group = skin_fixed_test_date_year), size = 1) +
  scale_color_viridis_c(
    option = "plasma",  # You can try "viridis", "inferno", "magma", etc.
    name = "Year"
  ) +
  labs(
    x = "Month",
    y = "Number of annual round tests (test type = 1)",
    color = "Year",
    title = "Monthly count of annual round tests (test type = 1) by year"
  ) +
  theme_jamie +
  theme_45_x +
  theme(plot.caption.position = "plot",
        #legend.title = element_text(size = 16, face = "bold"),
        plot.caption = element_text(size = 9, color = "black", face = "italic"),
        plot.title.position = "plot") +
  facet_wrap(~ herd_type_ml_description, scales = "free", ncol = 2)
```
:::

### All herds proportion

```{r}
#| label: fig-propsicttroundtest
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 11
#| fig.cap: "The proportion of all round tests by month"

round_tests <- master_tb %>% 
  filter(test_type == "1") %>% 
  mutate(skin_month = lubridate::month(skin_fixed_test_date, label = TRUE)) %>% 
  split(list(.$skin_fixed_test_date_year)) %>%
  map(~tabyl(.x, skin_month)) %>% 
  map(., ~mutate(.x, percent = round(percent, 2)) %>% select(-n)) %>% 
  imap(~mutate(.x, year = .y)) %>% 
  bind_rows() %>%
  mutate(year = as.numeric(year))
#round_tests


round_tests_month_year_plot <- ggplot(round_tests, aes(x = year, y = percent, fill = skin_month)) +
  geom_bar(position = position_fill(reverse = TRUE), stat = "identity", #color = 'black', 
           width = 0.9) +
  scale_fill_tableau20(name = "Month", direction = 1) +
  #scale_fill_viridis_d(name = "SICTT test-type", direction = 1) +
  theme_cowplot() +
  theme(axis.text = element_text(size = 12, face = "bold", colour = "black"),
        axis.title = element_text(size = 16, face = "bold", colour = "black"),
        #axis.text.x = element_blank(), #remove x axis title
        #axis.text.x = element_text(size = 12)
        #legend title only needed for first one as only extracting first
        legend.title = element_text(size = 10, face = "bold"), 
        legend.text = element_text(size = 10),
        #legend.position = "top",
        #grey background like here:
        #https://github.com/davidsjoberg/ggstream
        #plot.background = element_rect(fill = "grey88", color = NA),
        #make facet white rather than default grey
        #strip.background = element_rect(colour = "white", fill = "white")
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  ) + 
  background_grid() +
  labs(x = "Year", 
       y = "Proportion", 
       #fill = "AHCS herd type",
       title =
         paste0("Proportion of all round-test SICTT (test-type = 1) by year"),
       caption = "") +
  #theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  scale_x_continuous(breaks = years_for_study) +
  scale_y_continuous(labels = scales::label_percent(scale = 100), breaks = seq(0, 1, .1)) +
  coord_flip() +
  guides(x =  guide_axis(angle = 0)) +
  theme(plot.caption.position = "plot",
        #legend.title = element_text(size = 16, face = "bold"),
        plot.caption = element_text(size = 9, color = "black", face = "italic"),
        plot.title.position = "plot")

round_tests_month_year_plot
```

### by herd-type

```{r}
#| label: fig-propsicttroundtestherdtype
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 11
#| fig.cap: "The proportion of all round tests by month and herd-type"
#
round_tests_herd_type_month <- master_tb %>% 
  filter(test_type == "1") %>% 
  mutate(skin_month = lubridate::month(skin_fixed_test_date, label = TRUE)) %>% 
  filter(herd_type_ml_description  %in% c("Dairy", "Beef", "Fattener", "Mixed", "Store")) %>% 
  droplevels() %>% 
  split(list(.$herd_type_ml_description, .$skin_fixed_test_date_year)) %>%
  #map(., split(list(.$skin_fixed_test_date_year))) %>%
  
  map(~tabyl(.x, skin_month)) %>% 
  map(., ~mutate(.x, percent = round(percent, 2)) %>% select(-n)) %>% 
  imap(~mutate(.x, year = .y)) %>% 
  bind_rows() %>%
  mutate(herd_type_ml_description = str_extract(year, "[[:alpha:]]+"),
         year = as.numeric(str_extract(year, "\\d+")))
#round_tests_herd_type_month



round_tests_herd_type_month_plot_facet <- ggplot(round_tests_herd_type_month, aes(x = year, y = percent, fill = skin_month)) +
  geom_bar(position = position_fill(reverse = TRUE), stat = "identity", #color = 'black', 
           width = 0.9) +
  scale_fill_tableau20(name = "Month", direction = 1) +
  #scale_fill_viridis_d(name = "SICTT test-type", direction = 1) +
  theme_cowplot() +
  theme(axis.text = element_text(size = 12, face = "bold", colour = "black"),
        axis.title = element_text(size = 16, face = "bold", colour = "black"),
        #axis.text.x = element_blank(), #remove x axis title
        #axis.text.x = element_text(size = 12)
        #legend title only needed for first one as only extracting first
        legend.title = element_text(size = 10, face = "bold"), 
        legend.text = element_text(size = 10),
        #legend.position = "top",
        #grey background like here:
        #https://github.com/davidsjoberg/ggstream
        #plot.background = element_rect(fill = "grey88", color = NA),
        #make facet white rather than default grey
        #strip.background = element_rect(colour = "white", fill = "white")
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  ) + 
  background_grid() +
  labs(x = "Year", 
       y = "Proportion", 
       #fill = "AHCS herd type",
       title =
         paste0("Proportion of all round-test SICTT (test-type = 1) by year"),
       caption = "") +
  #theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  scale_x_continuous(breaks = years_for_study) +
  scale_y_continuous(labels = scales::label_percent(scale = 100), breaks = seq(0, 1, .1)) +
  coord_flip() +
  #guides(x =  guide_axis(angle = 0)) +
  theme_45_x +
  theme(plot.caption.position = "plot",
        #legend.title = element_text(size = 16, face = "bold"),
        plot.caption = element_text(size = 9, color = "black", face = "italic"),
        plot.title.position = "plot") +
  facet_grid(~herd_type_ml_description)
round_tests_herd_type_month_plot_facet

```

## APT (Animals Reactor per Thousand Animal Tests)

SICTT positive animals per thousand tested

APT is used by DAFM as an administrative metric (which just looks at SICTT)

```{r}
#| label: fig-aptskin
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 11
#| fig.cap: "APT"
# can we replicate Mikes APT plots
# this will just look at skin reactors so what I need is the number of animals tested
tmp <- master_tb %>% 
  select(herd_no, herd_type_ml_description, total_animals, total_reactor_skin, skin_fixed_test_date, skin_fixed_test_date_year) %>% 
  mutate(year_month = floor_date(skin_fixed_test_date , "month")) %>% 
  #mutate(skin_month = lubridate::month(skin_fixed_test_date, label = TRUE)) %>% 
  group_by(year_month) %>% 
  summarise(total_reactor_skin = sum(total_reactor_skin), total_animals_tested = sum(total_animals)) %>% 
  ungroup() %>% 
  mutate(apt_sictt_overall = (total_reactor_skin/total_animals_tested)*1000,
         year = year(year_month))

min_max_dates_df <- min_max_geom_react_with_var(tmp, apt_sictt_overall)
min_year <- min(master_tb$skin_fixed_test_date_year)
max_year <- min(master_tb$skin_fixed_test_date_year)
apt_plot_function_all(tmp, "year_month", "apt_sictt_overall")
```

### APT by herd

```{r}
#| label: fig-aptskinherd
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 11
#| fig.cap: "APT by herd-type"
tmp <- master_tb %>% 
  select(herd_no, herd_type_ml_description, total_animals, total_reactor_skin, skin_fixed_test_date, skin_fixed_test_date_year) %>% 
  mutate(year_month = floor_date(skin_fixed_test_date , "month")) %>% 
  #mutate(skin_month = lubridate::month(skin_fixed_test_date, label = TRUE)) %>% 
  group_by(year_month, herd_type_ml_description) %>% 
  summarise(total_reactor_skin = sum(total_reactor_skin), total_animals_tested = sum(total_animals)) %>% 
  ungroup() %>% 
  mutate(apt_sictt_overall = (total_reactor_skin/total_animals_tested)*1000,
         year = year(year_month)) %>% 
  filter(!herd_type_ml_description  %in% c("Trader", "Unknown", NA, "Store"))

min_max_dates_df <- min_max_geom_react_with_var(tmp, apt_sictt_overall)
#apt_plot_function(tmp, "year_month", "apt_sictt_overall")
apt_plot_function(tmp, "year_month", "apt_sictt_overall", facet_var = herd_type_ml_description)
```
