# Summary of main findings {#sec-summary}

------------------------------------------------------------------------

$~$ $~$

```{r}
#| label: global-options
#| include: false

# Center figures (alternative approach in Quarto)
knitr::opts_chunk$set(fig.align = "center")
# like this: #| fig-align: center
```

```{r}
#| label: readin-data
#| include: false
library(magick)

# demographics summary
demographics_summary <- image_read("figures/merged_plot_17_Dec_2025.png",
                                  density = 300) %>% 
  image_ggplot(interpolate = TRUE)

# cases summary
cases_summary <- image_read("figures/merged_plot_cases_prevalance_17_Dec_2025.png") %>% 
  image_ggplot(interpolate = TRUE)

# case distribution summary
bd_dist_summary <- image_read("figures/merged_plot_bd_size_cats_10_Sep_2025.png") %>% 
  image_ggplot(interpolate = TRUE)

# lesion summary
lesion_summary <- image_read("figures/lesion_plot_17_Dec_2025.png") %>% 
  image_ggplot(interpolate = TRUE)

# aggregated case count
nb_model_summary <- image_read("figures/negative_binomial_model_01_Oct_2025.png") %>% 
  image_ggplot(interpolate = TRUE)

```

The website contains a lot of output and so for the purposes of a quick summary, we have extracted a couple of plots from the whole website that summarise some of the main findings - these are some of the plots included in the journal submission.

# Demographics overview

```{r}
#| label: fig-demographics
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: "Three main demographic characteristics from cattle in Ireland: (a) proportion of herd-types, (b) proportion of the bovine population by herd-types and (c) the median herd size by herd-type, for each year 2008-2024."

demographics_summary
```

# bTB burden overview

```{r}
#| label: fig-cases
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: "Three plots illustrating the burden of bTB over time and the variation between different herd-types: (a) total bTB cases by diagnostic method, (b) proportion of all bTB cases by herd-type and (c) herd-level prevalence of herd types by year."

cases_summary
```

# bTB case distribution in herd-level BDs

```{r}
#| label: fig-casedistribution
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: "The distribution of the total number of bTB cases (via all diagnostic methods) in herd-level BDs in (a) continous form and (b) by categorizing them by different groupings of cases (1, 2-3,  4-5, 5+)."

bd_dist_summary
```

# Lesion overview

```{r}
#| label: fig-lesion
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: "Lesion prevalence among bTB cases (classified by SICTT or IFN-γ) and healthy animals sent to slaughter."

lesion_summary
```

# Aggregated case count model

```{r}
#| label: fig-aggcases
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: "Aggregated negative binomial regression model results exploring the association between year and herd-type on total count of bTB cases."

nb_model_summary
```
