# Time-series analysis and case forecasting {#sec-timeseries}

---

$~$ $~$


```{r}
#| label: global-options
#| include: false

# Center figures (alternative approach in Quarto)
knitr::opts_chunk$set(fig.align = "center")
# like this: #| fig-align: center
```

```{r}
#| label: readin-data
#| include: false
#setwd(rprojroot::find_rstudio_root_file())
# read_in_trends_short.R
#source(file.path("R_files", "read_in_quarto_trends.R"))

source("R_files/read_in_quarto_trends.R")
source("R_files/case_number_quarto_functions.R")
library(fpp3)

master_file <- dir_info("data", type = "file", regexp = "(?i)master.*\\.rds$") %>% 
  subset(subset = modification_time == max(modification_time), select = path, drop = T)
master_tb <- readRDS(master_file) %>% 
  filter(skin_fixed_test_date_year > 2007) %>% 
  filter(skin_fixed_test_date_year != 2025)


all_cases_collapsed_file <- dir_info("data", type = "file", regexp = "(?i)all_cases_collapsed.*\\.rds$") %>% 
  subset(subset = modification_time == max(modification_time), select = path, drop = T)
all_cases_collapsed <- readRDS(all_cases_collapsed_file) %>% 
     filter(year > 2007) %>% 
    filter(year != 2025)




min_year <- min(all_cases_collapsed$year)
max_year <- max(all_cases_collapsed$year)
years_for_study <- c(min_year:max_year)

case_count_month <- case_count_function(all_cases_collapsed, "year_month")
case_count_month <- case_count_month %>% 
  rename(year_month_date = year_month) %>% 
  mutate(year_month = yearmonth(year_month_character))

# just look at all cases for now
all_case_case_month <- case_count_month %>% 
  filter(best_estimate_gif_skin_lab_string == "All cases")

# set as tsibble:
all_case_case_month <- as_tsibble(all_case_case_month, 
                                  #key = best_estimate_gif_skin_lab_string, 
                                  index = year_month)

# read in decomp plot function:
source("R_files/time_series_decomp_plot_fun.R")

```


## bTB case count



### Seasonality


```{r}
#| label: fig-seasonplot
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 8 
#| fig.cap: "Seasonality of bTB cases by month."
seasonality_plot_btb_cases <- all_case_case_month %>%
  gg_subseries(cases) +
  labs(x = "Year",
       y = "No. of bTB cases",
       title =  paste0("Seasonality of bTB cases by month (", min_year, " - ", max_year, ")"),
       caption = paste0("The blue line indicates the average number of cases in each month", 
                        "\n", 
                        "bTB cases: any bTB case detected by any diagnostic method (SICTT, IFN-γ, culture positive)",
                        "\n",
                        "SICTT: Single Intradermal Comparative Tuberculin Test",
                        "\n",
                        "IFN-γ: interferon-gamma assay")) +
  theme_jamie +
  background_grid() +
  theme_90_x +
  scale_colour_nejm(name = "bTB cases") +
  scale_y_continuous(labels = scales::label_comma()) +
  #scale_x_type +
  #scale_x_continuous(breaks = nth_element(c(unique(df[[x_var]])), 1, 2)) +
  #scale_color_simpsons(name = "Quartile type") +
  # need this for bold headings
  theme(title = element_text(size = 18, face = "bold"),
        legend.text = element_markdown(size = 16),
        plot.caption.position = "plot",
        legend.title = element_text(size = 16, face = "bold"),
        plot.caption = element_text(size = 9, color = "black", face = "italic"),
        plot.title.position = "plot",
        # remove legend so its zoomed in, then add legend seperately
        legend.position = "top"
        #legend.position = c(0.2, 0.9))  
  ) 

seasonality_plot_btb_cases 
```


---


$~$ $~$

### STL decomposition

As an initial, exploratory step, we applied a seasonal-trend decomposition procedure based on loess (locally estimated scatterplot smoothing, a type of locally-weighted polynomial regression), abbreviated to STL [9]. The purpose of the decomposition of a time series is to “deconstruct” it into a number of component series which, when combined, will reproduce the original time series. Using STL, the specific aim is to split a seasonal time series into three additive components: the underlying trend, seasonality and the remainder (noise). 


```{r}
#| label: fig-stlplot
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 14
#| fig.cap: "STL decomposition of bTB cases by month."
stl_data <- all_case_case_month %>% 
  filter(year_month_date < as.Date("2025-01-01")) %>% 
  #     STL_model = STL(rate ~ trend() + season(window = "periodic"), weights = denominator)
  model(stl = STL(cases),
        #STL_model = STL(log(cases) ~ season(window = Inf))
  ) 

decomp_df <- components(stl_data) %>% 
  mutate(year = year(year_month),
         year_month_date = ymd(paste(year_month, "1"))) %>% 
  # messes up min_max_geom_react_with_var_STL and everything if you leave as tsibble
  as_tibble()

# essentially need to pivot this into long form
decomp_df_long <- decomp_df %>% 
  select(-.model) %>% 
  pivot_longer(
    cols = c(cases, trend, season_year, remainder, season_adjust),
    names_to = "component",
    #values_to = "Value"
  ) %>% 
  mutate(component = case_when(component == "cases" ~ "Cases",
                               component == "trend" ~ "Trend (loess smoother)",
                               component == "remainder" ~ "Remainder",
                               component == "season_year" ~ "Seasonal component",
                               component == "season_adjust" ~ "Seasonally adjusted"), 
         # If the seasonal component is removed from the original data, the resulting
         # values are the “seasonally adjusted” data.
         component = fct_relevel(component, c("Cases", 
                                              "Trend (loess smoother)",
                                              "Seasonal component",
                                              "Remainder",
                                              "Seasonally adjusted"))) %>% 
  filter(component != "Seasonally adjusted") %>% 
  # for now drop levels we dont need
  droplevels()


min_max_dates_df <- min_max_geom_react_with_var_STL(decomp_df, cases)
cases_decomp_plot <- time_series_month_plot_fun(decomp_df_long, "year_month_date", "value", month_interactive = FALSE, facet_var = component)
cases_decomp_plot
```


---


$~$ $~$

### Forecasting cases


```{r}
#| label: fig-forecastingcases
#| include: TRUE
#| echo: FALSE
#| eval: TRUE
#| warning: FALSE
#| message: FALSE
#| fig.width: 12
#| fig.height: 9
#| fig.cap: "Forecasting predicitions of monthly bTB cases for 2025."
train <- all_case_case_month %>% 
  filter(year_month_date < as.Date("2025-01-01"))

fit_models <- train %>% 
  model(
    arima = ARIMA(cases))

# edited forecast plot
min_max_geom_react_with_var <- function(df, variable) {
  
  warning("df has to be data.frame/tibble, not tsibble")
  
  df_new <- df %>% 
    mutate(min_value = min({{variable}}, na.rm = TRUE),
           max_value = max({{variable}}, na.rm = TRUE)) %>% 
    group_by(year) %>% 
    summarise(min_date = min(year_month_date),
              # add 1 year lubridate
              max_date = min_date %m+% years(1),
              # subtract 1 days
              max_date = max_date %m-% days(1),
              min_value = first(min_value),
              max_value = first(max_value)) %>% 
    ungroup() %>% 
    mutate(year_f = as.character(year))
  
}
min_max_dates_df <- min_max_geom_react_with_var(data.frame(all_case_case_month), cases) %>% 
  bind_rows(df <- data.frame(year = c(2025),
                             min_date = as.Date("2025-01-01"),
                             max_date = as.Date("2025-12-31"),
                             min_value = c(881),
                             max_value = c(5557),
                             year_f = as.character(2025)))
# do this first separately so that we can modify geom_layer
forecasts_df <- fit_models %>% 
  select(arima) %>% 
  forecast(h = 12) 
forecast_2025_plot <- forecasts_df %>% 
  autoplot(train) +
  geom_line(aes(year_month, y = .mean), alpha = 0.3, linewidth = 1.7, data = forecasts_df) +
  geom_line(aes(year_month, cases), linewidth = 1.2) +
  theme_jamie +
  scale_y_continuous(labels = scales::label_comma()) +
  scale_x_yearmonth(date_breaks = "1 year", date_labels = "%Y-%m-%d") +
  theme_45_x +
  geom_rect(
    data = min_max_dates_df, 
    mapping = aes(
      xmin = min_date,
      xmax = max_date,
      ymin = -Inf, # min_value
      ymax = Inf, # max_value,
      # ymin = min_value,
      # ymax = max_value,
      fill = year_f),
    inherit.aes = FALSE,
    #fill = df2$colour_me,
    alpha = 0.3) + 
  scale_fill_tableau20(name = "Year",
                       # turn off legend
                       guide = "none") +
  theme(plot.caption.position = "plot",
        plot.title.position = "plot") +
  labs(x = "Year",
       y = "bTB cases",
       title = "Monthly bTB cases and forecast predictions for 2025",
       caption = paste0("Predictions based of seasonal ARIMA time-series model (form ARIMA(0,1,1)(1,1,1)[12])",
                        "\n",
                        "bTB cases: any bTB case detected by any diagnostic method (SICTT, IFN-γ, culture positive)",
                        "\n",
                        "SICTT: Single Intradermal Comparative Tuberculin Test",
                        "\n",
                        "IFN-γ: interferon-gamma assay"),
       level = "Prediction\ninterval") +
  coord_cartesian(xlim = c(as.Date("2016-01-01"), as.Date("2026-01-01"))) 
forecast_2025_plot
```

